---
title: "DynamicStim"
author: "Sumner-Rooney et al."
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r include=FALSE}
library('tidyverse')
library('brms')
```

Add the data set.

```{r}
df_agg <- read_delim('DynamicStim.txt',delim = "\t",
                   col_types = "ffflllllnc")

df_agg %>% mutate(
  UnanT = if_else(Unanimous == "Y" & Viewer_1 == TRUE, TRUE, FALSE)) -> df_agg

head(df_agg)
```

### Aggregated observations

#### Skeleton model

Run a model with only the random intercepts and no main effect.

```{r}
prior <- c(prior(exponential(1), class="sd") )

m_agg_sk <- brm(UnanT ~ 1 + (1|Animal),
          family = bernoulli(), refresh = 0,
          prior = prior,
          data = df_agg)
m_agg_sk
```

```{r}
plot(m_agg_sk)
```

The data are mostly unsuccessful responses.

#### Effect of stimulus and individual (with random intercept)


```{r}
prior <- c(prior(normal(0,2), class="b"),
           prior(exponential(1), class="sd")
           )

m_agg <- brm(UnanT ~ Stimulus + (1|Animal),
          family = bernoulli(), refresh = 0,
          prior = prior,
          data = df_agg)
m_agg
```

```{r}
plot(m_agg)
```


```{r}
conditional_effects(m_agg)
```


```{r}
hyps <- c("StimulusL - Intercept > 0","StimulusO - Intercept > 0")

hyp <- hypothesis(m_agg,hyps)
hyp
```


```{r}
plot(hyp, ignore_prior = TRUE)
```


### Separate observations

```{r}
df_agg %>% 
  gather("Viewer","Response", Viewer_1:Viewer_5) %>%
  select(-Unanimous,-Response_rate,-UnanT) -> df_sep

head(df_sep)
```

#### Skeleton model

```{r}
prior <- c(prior(exponential(1), class="sd")    )

m_sep_sk <- brm(Response ~ 1 + (1|Animal) + (1|Viewer),
          family = bernoulli(), refresh = 0,
          prior = prior,
          data = df_sep)
m_sep_sk
```

```{r}
plot(m_sep_sk)
```


#### Main model

```{r}
prior <- c(prior(normal(0,2), class="b"),
           prior(exponential(1), class="sd")
          )

m_sep <- brm(Response ~ Stimulus + (1|Animal) + (1|Viewer),
          family = bernoulli(), refresh = 0,
          prior = prior,
          data = df_sep)
m_sep
```

```{r}
plot(m_sep)
```




```{r}
conditional_effects(m_sep)
```

Both of the experimental stimuli exceed the control. But by how much?

```{r}
hyps <- c("StimulusL - Intercept > 0","StimulusO - Intercept > 0")

hyp <- hypothesis(m_sep,hyps)
hyp
```


```{r}
plot(hyp, ignore_prior = TRUE)
```


Why does the separate viewer analysis differ from the aggregate? 

```{r}
df_sep %>% 
  group_by(Viewer) %>%
    summarise(proportion = sum(Response)) -> summation
summation
```

Similar numbers. Why did the skeleton model of the aggregate m_agg_sk suggest the mean response was below 50% if most responses are successes? Try again without the random effect.

```{r}
m_agg_1p <- brm(UnanT ~ 1,
          family = bernoulli(), refresh = 0,
          data = df_agg)
m_agg_1p
```


```{r}
sum(df_agg$UnanT)/length(df_agg$UnanT)
```
Because there are fewer - the unanimous criterion is too strict for the amount of data. Ff it required only a majority of successful observations from the Viewers per trial, most would be successful.

```{r}
sum(df_agg$Response_rate > 0.5) / length(df_agg$Response_rate)
```
#### Majoritarian

Let's run with that idea a minute

```{r}
df_agg %>%
  mutate(maj = if_else(Response_rate > 0.5, TRUE, FALSE)) -> df_agg
```




```{r}
prior <- c(prior(normal(0,2), class="b"),
           prior(exponential(1), class="sd")
           )

m_agg_maj <- brm(maj ~ Stimulus + (1|Animal),
          family = bernoulli(), refresh = 0,
          prior = prior,
          data = df_agg)
m_agg_maj
```

```{r}
plot(m_agg)
```



```{r}
conditional_effects(m_agg_maj)
```
Well now, still responds a bit when its the negative control. 

```{r}
df_agg$Response_rate[df_agg$Stimulus=="C"]
```
And the effect is much weaker in this case. 

```{r}
hyps <- c("StimulusL - Intercept > 0","StimulusO - Intercept > 0")

hyp <- hypothesis(m_agg_maj,hyps)
hyp
```

It is better to avoid dichotomizing data, where possible. This is clear in this case, where dichotomizing the observations by some number of *Viewers* reduces the effect size. I would proceed with the non-aggregated model: *m_sep*.




